---
title: "One-way ANOVA example - Crop yields"
output: 
  html_document: 
    theme: paper
    toc: yes
    toc_float: yes
  html_notebook: 
    toc: yes
    toc_float: yes
---

# Problem
(source: Crawley, The R book)

We have an experiment in which crop yields per unit area were measured from 10 randomly selected fields on each of three soil types. All fields were sown with the same variety of seed and provided with the same fertilizer and pest control inputs.

The question is whether soil type significantly affects crop yield, and if so, to what extent.

```{r}
sand <- c(6,10,8,6,14,17,9,11,7,11)
clay <- c(17,15,3,11,14,12,12,8,10,13)
loam <- c(13,16,9,12,15,16,17,13,18,14)
dati <- data.frame(sand,clay,loam)
dati
```


# Tasks

* Calculate the mean for each soil and make a boxplot for the data
* The data format is not appropriate for subsequent analysis, prepare the data by creating a new data frame with a column for the soil type!
* One of the assumptions of ANOVA is about the homogeneity of variances - check for heteroschedasticity!
* Continue with the analysis of variance
* Graphically check for the assumptions
* Fit a linear model to the data to find out between which soil types the differences are actually significant


## Calculate the mean for each soil and make a boxplot for the data

```{r}
colMeans(dati)
boxplot(dati)
```

## Change the data format (da "wide data" a "long data")

```{r}
soil.type <- c(rep("sand",10),
               rep("clay",10),
               rep("loam",10))
yields <- c(sand, clay, loam)
dati.new <- data.frame(soil.type, yields)
dati.new
```

## Check for heteroschedasticity

```{r}
by(dati.new$yields, dati.new$soil.type, var) # restituisce le varianze per ogni soil.type
fligner.test(yields ~ soil.type, data=dati.new) # test specifico che verifica omogeneità
```

There is *no evidence of any significant difference* in variance across the three samples, so it is legitimate to continue with the analysis. Quindi va bene considerare le varianze omogenee.

## Continue with the analysis of variance

```{r}
anova.yields <- aov(yields ~ soil.type, data=dati.new)
summary(anova.yields)
```

C'è almeno un tipo di suolo con resa diversa rispetto gli altri due.

```{r}
anova(lm(yields ~ soil.type, data=dati.new)) 
# fa la stessa cosa di prima solo un altro modo di scriverlo
```

## Graphically check for the assumptions

```{r}
anova.yields <- aov(yields ~ soil.type, data=dati.new)
plot(anova.yields) # stessi grafici che si ottengono da un modello di regressione lineare
# il Q-Q plot ci assicura che i residui della varianza hanno distribuzione normale
```

The first plot check the most important assumption of constancy of variance: there should be no pattern in the residuals against the fitted values, and, indeed, there is none.   
The second plot tests the assumption of normality of errors: there should be a straight-line relationship between our standardized residuals and theoretical quantiles derived from a normal distribution. Points 6, 11 and 13 lie a little off the straight line, but this is nothing to worry about.   
The third plot shows that the residuals are well behaved.   
The fourth plot shows that there are no highly influential values that might be distorting the parameter estimates. 

## Fit a linear model to the data to find out between which soil types the differences are actually significant, due modi per procedere

```{r}
mod <- lm(yields ~ soil.type, data=dati.new)
summary(mod)
```

soil.type clay è stato inglobato nell'intercetta e diventa quindi il soil.type contro il quale vengono confrontati sia sand che loam, Rprende come riferimento sempre il PRIMO IN ORDINE ALFABETICO.
l'output ci dice che tra sand e clay non si trova differenza significativa sulla resa, invece si trova una differenza QUASI significativa tra loam e clay.

```{r}
dati.relevel <- dati.new
dati.relevel$soil.type <- relevel(as.factor(dati.relevel$soil.type), ref="sand") 
# si riordano i dati per usare sand come riferimento per il confronto, bisogna usare as.factor() per rendere noto ad R di trattare soil.type come un fattore e non delle stringhe

mod.relevel <- lm(yields ~ soil.type, data=dati.relevel)
summary(mod.relevel)
```

Questo test di Tuckey permette di verificare tutti i possibili confronti a coppie e restituisce quali sono significativi, calcolando le differenze tra le rese.
Ci dice la differenza, l'intervallo di confidenza e un p-value aggiustato (si usa sempre il cut off 0.05), nel background in realtà il valore soglia in questo caso dovrebbe essere 0.05 diviso il numero di possibili confronti a coppie (in questo caso 3), questa severità è necessaria perchè man mano che si testano possibili coppie diventa più probabile trovare p-value sotto 0.05
*Per questo il p-value è "aggiustato"*

```{r}
# Tukey Honestly Significant Differences
TukeyHSD(anova.yields) # where anova.yields comes from aov()
plot(TukeyHSD(anova.yields)) 
# nota che solo la coppia sand-loam non interseca lo zero (statisticamente significativo)
```

Il test di bonferroni è il metodo più severo possiibile per cercare significatività a coppie. *Il p-value è già aggiustato su un valore soglia ancora minore rispetto a Tuckey*

In generale di solito è sufficiente il test di Tuckey, bonferroni è un po' overkill a meno che non si stia difendendo una teoria controversa.

```{r}
pairwise.t.test(dati.new$yields, dati.new$soil.type, p.adjust.method = "bonferroni")
```

