---
title: "Snow Vole Captures - ggplot2"
output:
  word_document: 
    toc: yes
    fig_caption: yes
  html_document:
    df_print: paged
---

Dataset: Snow Vole Captures (Chionomys nivalis)

```{r}
captures <- read.csv("data/captures.csv", sep=";")

captures <- subset(captures, sex %in% c("M", "F"))

library("ggplot2")
```

Objective: Use the ggplot2 package to explore the morphological traits and population structure of the snow vole.

Task 1: Weight vs. Foot Length (Scatter Plot).
Investigate if heavier individuals have longer feet.

```{r}
t1 = ggplot(captures, aes(x = weight_g, y = footlength_mm, color = sex)) + geom_point(na.rm = TRUE) + geom_smooth(se=FALSE, method = "lm", na.rm = TRUE) + ggtitle("Relationship Between Weight and Foot Length") + xlab("Weight [g]") + ylab("Foot Length [mm]")

ggsave("output/weight_footlength_scatter.png", t1)

t1
```

The data shows a positive relationship between the foot length and the weight of the captured Snow Voles, this goes for males and females both.

Task 2: Weight Distribution (Histogram).
Analyze how weight is distributed across the population.

```{r}
t2 = ggplot(captures, aes(x = weight_g)) + geom_histogram(na.rm = TRUE, binwidth = 2) + facet_wrap("session", scale = "free_y") + ggtitle("Weight Distribution In Each Capture Session") + xlab("Weight [g]") + ylab("Frequency")

ggsave("output/weight_hist.png", t2)

t2
```

Task 3: Comparing Sex and Age (Boxplot).
Check for sexual dimorphism or differences between adults and juveniles.

```{r}
t3 = ggplot(captures, aes(x = age, y = weight_g, fill = sex)) + geom_boxplot(na.rm = TRUE) + ggtitle("Sexual Dimorphism Across Age Groups") + xlab("Age") + ylab("Weight [g]")

ggsave("output/weight_age_box.png", t3)

t3
```

The heaviest group is the one for adult males. Some outliers in the juvenile group, also interesting the fact that there's only one confirmed juvenile female captured.
