---
title: "One sample t-test example"
output: html_notebook
---

# Problem

Consider the variable **rat**. It is included in the **UsingR** package so you have to load it first.

* Describe and plot its distributions, then assess departure from normality using an appropriate test and plot.
* Based on the results of the normality test, could you perform a t-test to assess whether the sample mean is significantly larger than 110?
* Comment on the results and compare them to the results of the corresponding non-parametric test.


# Solution

```{r package, message=FALSE}
#install.packages("UsingR")
library(UsingR)
```


```{r}
# load the data
data(rat)
```

```{r}
# histogram
hist(rat)
```

```{r}
# stripcharts
stripchart(rat)
stripchart(rat, method = "stack")
stripchart(rat, method = "jitter")
```

```{r}
# boxplot
boxplot(rat)
```

```{r}
# normality test
shapiro.test(rat)
qqnorm(rat); qqline(rat)
```



```{r}
t.test(rat, mu=110, alternative="greater")
```



```{r}
wilcox.test(rat, mu=110, alternative="greater")
```