---
title: "Two sample t-test example #2"
output: html_notebook
---

# Problem - dilemma of two graders

In order to promote fairness in grading, each application was graded twice by different graders. Based on the
grades, can we see if there is a difference between the two graders? The data is

```{r}
x <- c(3, 0, 5, 2, 5, 5, 5, 4, 4, 5)
y <- c(2, 1, 4, 1, 4, 3, 3, 2, 3, 5)
```

Clearly there are differences. Are they described by random fluctuations, or is there a bias of one grader over another? 
Verify the t-test assumptions, and then proceed with the t-test if appropriate. Which is the corresponding non-parametric test? Perform it and compare the results to the one obtained with the t-test.

# Solution

```{r}
# normality test
par(mfrow=c(1,2))
qqnorm(x); qqline(x, col="red")
qqnorm(y); qqline(y, col="red")
shapiro.test(x)
shapiro.test(y)
```


```{r}
# homogeneity of variances
boxplot(x,y)
var.test(x,y)
```

```{r}
t.test(x,y, paired = TRUE,
       var.equal = TRUE)
```

```{r}
wilcox.test(x,y, paired = TRUE)
```