---
title: "Chisquare exercises"
output: html_notebook
---

# Code for slide examples

Remember that `df = (nrows-1)(ncols-1)`.

```{r gof-test}
## chi2 - goodness of fit test
freq <- c(22, 21, 22, 27, 22, 36)
probs <- rep(1/6, 6)
chisq.test(freq, p = probs)
```

```{r independence-test}
## chi2 - test of independence
yesbelt <- c(12813, 647, 359, 42)
nobelt <- c(65963, 4000, 2642, 303)

chisq.test(data.frame(yesbelt, nobelt))

names(chisq.test(data.frame(yesbelt, nobelt)))

data.frame(yesbelt, nobelt)
rowSums(data.frame(yesbelt, nobelt))
colSums(data.frame(yesbelt, nobelt))

# for the first cell
78776*13861/sum(data.frame(yesbelt, nobelt))

chisq.test(data.frame(yesbelt, nobelt))$exp

a <- chisq.test(data.frame(yesbelt, nobelt))

a
names(a)
a$p.value
```


```{r homogeneity-test}
## chi2 - test of homogeneity
die.fair <- sample(1:6, 200, p=c(1,1,1,1,1,1)/6, replace = TRUE)
die.bias <- sample(1:6, 200, p=c(0.5,0.5,1,1,1,2)/6, replace = TRUE)

res.fair <- table(die.fair)
res.bias <- table(die.bias)

rbind(res.fair, res.bias)

chisq.test(rbind(res.fair, res.bias))

names(chisq.test(rbind(res.fair, res.bias)))
```

# Exercices

## Exercise 1

A fish survey is done to see if the proportion of fish types is consistent with previous years. Suppose, the 3 types of fish recorded: parrotfish, grouper, tang are historically in a 5:3:4 proportion and in a survey the following counts are found:

```{r}
freq <- c(53, 22, 49)
probs <- c(5/12, 3/12, 4/12)
chisq.test(freq, p = probs)
```


## Exercise 2

In an effort to increase student retention, many colleges have tried block programs. Suppose 100 students are broken into two groups of 50 at random. One half are in a block program, the other half not. The number of years in attendance is then measured. We wish to test if the block program makes a difference in retention.
The data is:

```{r}
nonblock <- c(18, 15, 5, 8, 4)
block <- c(10, 5, 7, 18, 10)
chisq.test(data.frame(nonblock, block))
```

What chi-square test is appropriate here? Does the block program actually make a difference?

## Exercise 3

A survey of drivers was taken to see if they had been in an accident during the previous year, and if so was it a minor or major accident. The results are tabulated by age group:

```{r}
none <- c(67, 42, 75, 56, 57)
minor <- c(10, 6, 8, 4, 15)
major <-  c(5, 5, 4, 6, 1)
names <- c("under 18", "18-25", "26-40", "40-65", "over 65")
data.frame(names, none, minor, major)
```

Do a chi-squared hypothesis test of homogeneity to see if there is difference in distributions based on age.
