---
title: "Exercise - analysis of covariance (ANCOVA)"
output: 
  html_notebook: 
    toc: yes
    toc_float: yes
  html_document: 
    theme: paper
    toc: yes
    toc_float: yes
---

(Source: Crawley, The R book)

# Definition

Analysis of covariance (ANCOVA) combines elements from regression and analysis of
variance. The response variable is continuous, and there is at least one continuous explanatory
variable and at least one categorical explanatory variable.

# Example

Our worked example concerns an experiment on the
impact of grazing on the seed production of a biennial plant.    
Forty plants were allocated to
two treatments, grazed and ungrazed, and the grazed plants were exposed to rabbits during
the first two weeks of stem elongation. They were then protected from subsequent grazing
by the erection of a fence and allowed to regrow. Because initial plant size was thought
likely to influence fruit production, the diameter of the top of the rootstock was measured
before each plant was potted up. At the end of the growing season, the fruit production (dry
weight in milligrams) was recorded on each of the 40 plants, and this forms the response
variable in the following analysis.   

The object of the exercise is to estimate the parameters of the minimal adequate model
for these data.


```{r}
regrowth <- read.table("data/ipomopsis.txt", sep="\t", header=T)
head(regrowth)
```

* Describe the dataset and look at the distribution of the dependent variable.
* Inspect the data plotting fruit production against root size for each of the two treatments separately
* Calculate the mean fruit production for grazed and ungrazed plants and plot the data: are the means statistically different? Specify the null and alternative hypothesis and test using a t-test
* Now fit a regression model including both Root and Grazing as explanatory variables and comment on the results.
* Try to remove the interaction terms and compare the first and second model through an ANOVA table.
* Further simplify the model removing Grazing and compare again through an ANOVA table.
* Compare the models using the *step* function.

## Describe the dataset and look at the distribution of the dependent variable

```{r}
str(regrowth)
```

```{r}
hist(regrowth$Fruit)
shapiro.test(regrowth$Fruit)
qqnorm(regrowth$Fruit);qqline(regrowth$Fruit)
```

```{r}
summary(regrowth)
```



## Inspect the data plotting fruit production against root size for each of the two treatments separately

```{r}
grazed <- subset(regrowth, Grazing == "Grazed")
ungrazed <- subset(regrowth, Grazing == "Ungrazed")
with(grazed,
     plot(Fruit ~ Root, pch=16, col="red", xlim=c(4,11), ylim=c(14,117))) # look at the summary above to set the limits
with(ungrazed,
     points(Fruit ~ Root, pch=17, col="grey"))
```

The root size of plants in grazed plots is larger!

## Calculate the mean fruit production for grazed and ungrazed plants and plot the data: are the means statistically different? Specify the null and alternative hypothesis and test using a t-test

```{r}
mean(grazed$Fruit)
mean(ungrazed$Fruit)
# or
library(dplyr)
regrowth %>% 
  group_by(Grazing) %>% 
  summarize(mean(Fruit))
```

```{r}
with(regrowth,
     boxplot(Fruit ~ Grazing))
```

```{r}
var(grazed$Fruit, na.rm = T)
var(ungrazed$Fruit, na.rm = T)
var.test(grazed$Fruit,ungrazed$Fruit)
```



```{r}
with(regrowth,
     t.test(Fruit ~ Grazing))
```

Grazing seems to increase fruit production!

## Now fit a regression model including both Root and Grazing as explanatory variables and comment on the results

```{r}
ancova <- lm(Fruit ~ Grazing * Root, data = regrowth)
summary(ancova)
```

This shows that initial root size has a massive effect on fruit production (t =15.182), but
there is no indication of any difference in the slope of this relationship between the two
grazing treatments (this is the Grazing by Root interaction with t =0.321). The
ANOVA table for the maximal model looks like this:

```{r}
anova(ancova)
```


## Try to remove the interaction terms and compare the first and second model through an ANOVA table

```{r}
ancova2 <- lm(Fruit ~ Grazing + Root, data = regrowth)
summary(ancova2)
```

```{r}
anova(ancova, ancova2)
```

This says that model simplification was justified because it caused a negligible reduction in
the explanatory power of the model (p=0.75; to retain the interaction term in the model
we would need p<0.05).

## Further simplify the model removing Grazing and compare again through an ANOVA table

```{r}
ancova3 <- lm(Fruit ~ Root, data = regrowth)
summary(ancova3)
```


```{r}
anova(ancova2, ancova3)
```

This model simplification is a step too far. Removing the Grazing term causes a massive
reduction in the explanatory power of the model, with an F value of 115.63 and a vanishingly
small p value. The effect of grazing in reducing fruit production is highly significant and
needs to be retained in the model.

## Compare the models using the *step* function

This function causes all the terms to be tested to see whether they are needed in the minimal
adequate model. The criterion used is AIC, Akaike’s information criterion. In the
jargon, this is a ‘penalized log-likelihood’. What this means in simple terms is that it weighs
up the inevitable trade-off between degrees of freedom and fit of the model.

```{r}
step(ancova)
```


# Summary

In the analyses of covariance, first you compute two different regression lines.   
Next, you see whether the slopes are significantly different from zero.

* If the slopes of the regression lines are different, the lines cross each other somewhere (you check for the significance of the interaction term). One group has higher Y values in one part of the graph and lower Y values in another part of the graph. In this case we are thus not very interested in the difference between the intercepts.  
* If the slopes are not significantly different, you then draw a regression line through each group of points, all with the same slope.   
Then, you test the null hypothesis that all of the Y intercepts of the regression lines with a common slope are the same. Because the lines are parallel, saying that they are significantly different at one point (the Y intercept) means that the lines are different at any point.